---
title: "PRS_genepy"
author: "Emad"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document: 
    toc: true
    toc_float: true
    number_sections: true

---

```{r setup, vinclude=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, error = FALSE, message = FALSE, options(digits = 3),comment = FALSE )
```

```{r load_libraries, include=FALSE}
# Organize uk_biobank phenotype data ----

# 1.0 Load libraries ----
print("install and loading dependencies packages")

installed.packages("pacman") 
library("pacman")

#devtools::install_github("kenhanscombe/ukbtools", build_vignettes = TRUE, dependencies = TRUE, force = TRUE)

pacman::p_load(tidyverse     # tidy daya
               , ukbtools    # workhorse package for organizing ukbb
               , tictoc      # checking time 
               , furrr       # functional programing with map()
               , future      # parallel computing
               , arrow       # apache arrow provides powerful package for sharing data
               , data.table  # exporting data as .txt
               , ggpubr      # add themes and visualization prop. to ggplot
               ,plotly       # interactive plots
               ,caret        # partition data
               ,rcompanion   # calculating R2
               ,pROC         # calculating and plot AUC 
               ,plotROC      # calculating and plot ggplot
               ,gridExtra    # add table to ggplot
               )    
```

```{r importing files, include=FALSE}

print("source change_me.r file")
source(here::here("phenotype","organize_ukbb","change_me.R"))
source(here::here("phenotype","organize_ukbb","test.R"))
source(here::here("phenotype","organize_ukbb","organize_ukbb.R"))
source(here::here("PRS","import_prs_scores.R"))

```

# BOADICEA (313 Snps) {.tabset}

## all cases

```{r boadicea_all}
predict_auc(all_scores_sum_313)
```


## only prevalent

```{r boadicea_prev}
predict_auc(prev_scores_sum_313)
```

## only incident

```{r boadicea_inci}
predict_auc(inc_scores_sum_313)
```

# Khera (5234 Snps) {.tabset}

## all cases

```{r khera_all}
predict_auc(all_scores_sum_khera)
```


## only prevalent 

```{r khera_prev}
predict_auc(prev_scores_sum_khera)
```

## only incident

```{r khera_inci}
predict_auc(inc_scores_sum_khera)
```


# Breast Cancer Association Consortium (BCAC)  {.tabset}

## all cases 

```{r bcac_all}
predict_auc(all_scores_sum_bcac)
```


## only prevalent

```{r bcac_prev}
predict_auc(prev_scores_sum_bcac)
```

## only incident 

```{r bcac_inci}
predict_auc(inc_scores_sum_bcac)
```

# Genepy {.tabset}

## Breast Cancer {.tabset}

### AUC with age and PC 

```{r integrate_gps_prs}
integrate_gps_prs(training,testing)

```


### AUC without age

```{r breast_gps_prs}

gps_prs(training,testing)

```




## CAD {.tabset}

### AUC with age, sex and PC 

```{r}
integrate_cad_gps_prs(training_cad,testing_cad)
```

### AUC without age and sex 

```{r}

cad_gps_prs(training_cad,testing_cad)

```


## Diabetes II {.tabset}

### AUC with age, sex and PC

```{r}
integrate_diabetes_gps_prs(training_diabetes,testing_diabetes)
```


### AUC without age and sex 

```{r}
diabetes_gps_prs(training_diabetes,testing_diabetes)
```
