---
title: "Explore UK Biobank data"
author: "Emad"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document: 
    toc: true
    toc_float: true
    number_sections: true

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, error = FALSE, message = FALSE, options(digits = 3), comment = NA)

library("pacman")

#devtools::install_github("kenhanscombe/ukbtools", build_vignettes = TRUE, dependencies = TRUE, force = TRUE)

pacman::p_load(tidyverse     # tidy daya
               , ukbtools    # workhorse package for organizing ukbb
               , tictoc      # checking time 
               , furrr       # functional programing with map()
               , future      # parallel computing
               , arrow       # apache arrow provides powerful package for sharing data
               , data.table  # exporting data as .txt
               , ggpubr      # add themes and visualization prop. to ggplot
               , plotly      # interactive plots
               )  
```

```{r load}
source("/home/opc/projects/ukbiobank_prs/phenotype/organize_ukbb/change_me.R")
source("/home/opc/projects/ukbiobank_prs/phenotype/organize_ukbb/test.R")
source("/home/opc/projects/ukbiobank_prs/phenotype/organize_ukbb/organize_ukbb.R")
```
# Demographics {.tabset}

## ethinc background

```{r}
p <- ukbb_df %>% 
   count(ethnic_background_f21000_0_0) %>% 
 ggplot(aes(x=ethnic_background_f21000_0_0, y = n, fill = ethnic_background_f21000_0_0))+
   geom_bar(stat = "identity", position = "dodge", width = 0.75) +
   coord_flip()+
   theme(axis.text.y = element_text(angle = 60, hjust = 1))+
   labs(x= "count")+
   theme_pubr(legend = "none")+
   theme(axis.title.y = element_blank())

options(scipen=500000)
ggplotly(p)
```

## sex

```{r}

p <- ukbb_df %>% 
   count(sex_f31_0_0) %>% 
   ggplot(aes(x=sex_f31_0_0, y = n, fill = sex_f31_0_0))+
   geom_bar(stat = "identity", position = "dodge", width = 0.75) +
   scale_fill_manual( "legend", values = c("Female" = "pink", "Male" = "grey"))+
   coord_flip()+
   theme(axis.text.y = element_text(angle = 60, hjust = 1))+
   labs(x= "count")+
   theme_pubr(legend = "none")+
   theme(axis.title.y = element_blank())

options(scipen=500000)
ggplotly(p)


```

## age distribution

```{r}
p <- ukbb_df %>% 
   ggplot(aes(x=age_when_attended_assessment_centre_f21003_0_0, fill = sex_f31_0_0)) + 
   geom_density(na.rm = TRUE,alpha=0.3) +
 #  scale_fill_manual(values = c("grey35")) +
   scale_fill_manual(values = c( "hotpink",  "grey")) +
   theme(legend.position = "none",
         axis.title.x = element_text(face = "bold"),
         panel.grid = element_blank()) +
   labs(x = "Age") +
   guides(fill=guide_legend(title="sex"))+
   theme_pubr()

ggplotly(p)
```


# Demographics of a UKB sample subset (eg BMI>25)

As a quick qc step, want to look at the demographics of a particular subset of the UKB sample (BMI>25) relative to a reference sample (BMI<25).

```{r demographics,, fig.height=15, fig.width=15}
#ukbb_df<- ukbb_df 


subgroup_of_interest <- (ukbb_df$body_mass_index_bmi_f21001_0_0 >= 25)

ukb_context(ukbb_df , subset.var = subgroup_of_interest)

```

# Frequency of an ICD diagnosis by a target variable

## Gender

shows frequency of the query diagnoses with respect to gender

```{r gender ,out.width = '100%'}

ukb_icd_freq_by(ukbb_df , reference.var = "sex_f31_0_0", 
  icd.code = c("^(I2[0-5])", "^(I6[0-9])", "^(J09|J1[0-9]|J2[0-2]|P23|U04)", "^(C50[0-9])"),
  icd.labels = c("coronary artery disease", "cerebrovascular disease",
    "lower respiratory tract infection", "breast cancer") ,
  freq.plot = FALSE) %>%
  select(-dx) %>%  
  tidyr::gather(key = "disease", value = "frequency", -categorized_var) %>% 
  ggplot2::ggplot(aes(categorized_var, frequency, group = disease,
                      fill = disease)) +
  labs(x = "", y = "UKB disease frequency", color = "", fill = "",
       title = "") +
  # theme(title = element_text(face = "bold"), panel.grid = element_blank(),
  #       panel.background = element_rect(color = NULL,
  #                                       fill = alpha("grey", 0.10)),
  #       legend.key = element_blank(), axis.ticks.x = element_blank()) +
  scale_y_continuous(labels = scales::percent_format(2))+
  geom_bar(stat = "identity", position = "dodge",  width=0.7) +
  guides(fill = guide_legend(ncol = 1), size = FALSE,
         color = FALSE) +
  theme_pubr()

```

## Body mass index (BMI)

shows frequency of the query diagnoses with respect to BMI

```{r bmi ,out.width = '100%'}
ukb_icd_freq_by(ukbb_df , reference.var = "body_mass_index_bmi_f21001_0_0",
  icd.code = c("^(I2[0-5])", "^(I6[0-9])", "^(J09|J1[0-9]|J2[0-2]|P23|U04)", "^(C50[0-9])"),
  icd.labels = c("coronary artery disease", "cerebrovascular disease",
    "lower respiratory tract infection", "breast cancer") ,
  freq.plot = FALSE) %>% 
  select(-dx) %>% 
  dplyr::mutate(mid = (lower + upper) / 2) %>%
  tidyr::gather(key = "disease", value = "frequency", -categorized_var,
                -lower, -upper, -mid) %>%
  ggplot2::ggplot(aes(mid, frequency, group = disease, color = disease)) +
  labs(x = "Body mass index (BMI)", y = "UKB disease frequency", color = "", fill = "",
       title = "")  +
  # theme(title = element_text(face = "bold"), panel.grid = element_blank(),
  #       panel.background = element_rect(color = NULL,
  #                                       fill = alpha("grey", 0.10)),
  #       legend.key = element_blank(), axis.ticks.x = element_blank()) +
  scale_y_continuous(labels = scales::percent_format(2)) +
  geom_point(size = 2) +
  geom_line(size = 0.5) +
  guides(color = guide_legend(ncol = 1), size = FALSE,
         fill = FALSE) +
  theme_pubr()

```


```{r,out.width = '100%', eval=FALSE}
ukbb_df%>% 
#  summarise( Subjects = n() , Male =  sum(sex_f31_0_0 == "Male")/Subjects , 
#             Female = sum(sex_f31_0_0 == "Female")/Subjects) %>%
#  select(-Subjects) %>% 
#  gather(sex, percentage) %>% 
  ggplot(aes(x = sex_f31_0_0,fill=sex_f31_0_0 )) +
  geom_bar(aes(y = (..count..)/sum(..count..)),  width=0.4) +
  geom_text(aes(y = ((..count..)/sum(..count..)), label = scales::percent((..count..)/sum(..count..))),
            stat = "count", vjust = -0.25)  +
  coord_flip() +
  scale_y_continuous(labels = scales::percent, limits = c(0,1)) +
  labs(x = "", y = "", color = "", fill = "",  title = "") +
  theme_pubr()
  theme(title = element_text(face = "bold"), panel.grid = element_blank(),
        panel.background = element_rect(color = NULL,
                                        fill = alpha("grey", 0.10)),
        legend.key = element_blank(), axis.ticks.x = element_blank()) 

```


